Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-07-10 Report/joint_wsrm_formulation.tex
modified for KKT MMSE

The queue minimizing algorithms are discussed extensively in the networking literature for providing congestion-free routing between any two nodes\footnote{refers to routers or PC's} in the network. It is achieved by finding a control policy considering the differential backlogs of the packets present in the source and the destination node, where control policy refers to the resource or rate allocation strategy. The control policy satisfying the condition is known as \emph{backpreassure algorithm}, discussed extensively in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm assumes the physical layer rate as a fixed entity due to the wired infrastructure. In order to use the algorithm in the wireless network, the user rate variable \me{t_{k}} should be modeled in accordance with the wireless backbone network.

The \ac{Q-WSRM} scheme is formulated by extending the \emph{backpreassure algorithm} for the \ac{MIMO} \ac{OFDM} network, where multiple \acp{BS} acts like a source nodes and the user terminals as a receiver nodes. The user rate variables \me{t_k} are obtained by the design of the transmit and the receive precoders to maximize the weighted sum rate. The weights used in the precoder design algorithm should be a function of the number of backlogged packets corresponding to each user. In order find the weights, following the approach in \cite{neely2010stochastic}, we bound the Lyapunov function\footnote{\me{\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q^2_k(i)}} drift as
\begin{IEEEeqnarray}{rCl}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }, \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot and the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
Now the conditional Lyapunov drift for the instant \me{i} is given by
\begin{IEEEeqnarray}{rCl}
\Delta(\mbf{Q}(i)) &\triangleq& \mathbb{E} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \mathbb{E} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) + \mathbb{E}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)}, \IEEEyessubnumber
\end{IEEEeqnarray}
where the expectation depends on the control decisions made for the transmission variables \me{t_k} and the second term on the r.h.s is due to i.i.d assumption on the arrivals. We can bound the first term on the r.h.s by a constant \me{B} for all possible control actions taken for a given channel condition \cite{neely2010stochastic}, the resulting problem can be given as
\begin{equation}
\underset{t_k}{\text{maximize}} \quad \sum_{k \in \mc{U}} \, Q_k(i) \, t_k(i),
\label{backpreassure}
\end{equation}
where the expectation operator is dropped since it is maximized by maximizing the function inside in an opportunistic manner.

>>>>

The queue minimizing algorithms are discussed extensively in the networking literature for providing congestion-free routing between any two nodes\footnote{routers or user terminals} in the network. It is achieved by finding a control policy considering the differential backlogs of the packets present in the source and the destination node, where control policy refers to the resource or rate allocation strategy. The control policy satisfying the condition is known as \emph{backpreassure algorithm}, discussed extensively in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm assumes the physical layer rate as a fixed entity due to the wired infrastructure. In order to use the algorithm in the wireless network, the user rate variable \me{t_{k}} should be modeled in accordance with the wireless backbone network.

The \ac{Q-WSRM} scheme is formulated by extending the \emph{backpreassure algorithm} for the \ac{MIMO} \ac{OFDM} network, where multiple \acp{BS} act like a source nodes and the user terminals as a receiver nodes. The user rate variables \me{t_k} are obtained by the design of the transmit and the receive precoders to maximize the weighted sum rate. The weights used in the precoder design algorithm should be a function of the number of backlogged packets corresponding to each user. In order to find the weights, we adopt the Lyapunov drift minimization as discussed in \cite{neely2010stochastic}, where the Lyapunov function is given by the squared sum of the number backlogged packets as
\begin{equation}
\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k(i),
\end{equation}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot. The Lyapunov function provides a measure of congestion in the system, as discussed in \cite[Ch. 3]{neely2010stochastic}. Now the Lyapunov function drift is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }, \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
In order to minimize the number of backlogged packets at each instant, minimization of the Lyapunov drift \eqref{eqn-3.1} is carried over all possible control decisions in the form of transmission rates \me{t_k} to users in the system. The Lyapunov drift conditioned on the current backlogged packets \me{\mbfa{Q}(i)} is given by
\begin{IEEEeqnarray}{rCl}
\underset{\mbf{t}}{\text{minimize}} && \Delta(\mbf{Q}(i)) \triangleq \mathbb{E} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \mathbb{E} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) + \mathbb{E}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)}, \IEEEyessubnumber
\end{IEEEeqnarray}
where the second term on the r.h.s is due to i.i.d assumption on the arrivals. We can bound the first term on the r.h.s by a constant \me{B} for all possible control actions taken for a given channel condition \cite{neely2010stochastic}, the resulting problem can be given as
\begin{equation}
\underset{t_k}{\text{maximize}} \quad \sum_{k \in \mc{U}} \, Q_k(i) \, t_k(i),
\label{backpreassure}
\end{equation}
where the expectation operator is dropped since it is maximized by maximizing the function inside in an opportunistic manner.

### 2014-07-10 Report/kkt_mse_approach.tex
modified for KKT MMSE

By removing the absolute value operator from \eqref{eqn-mse-2}, the problem can be rewritten as
\begin{IEEEeqnarray}{rCl}\label{kkt-mse-1}
\underset{\substack{t_{l,k,n},\mvec{M}{k,n}, \alpha_{l,k,n}, \\ \delta_{b},\sigma_{l,k,n},\mvec{W}{k,n}}}{\text{minimize}} &\hspace{0.5cm}& \sum_{b \in \mc{B}} \sum_{k \in \mc{U}_b} \, a_k \, \left ( Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} t_{l,k,n} \right )^q \IEEEyessubnumber \label{kkt-mse-1.1} \\
\text{subject to} & \hspace{0.5cm} & \nonumber \\
\alpha_{l,k,n} : & \hspace{0.5cm} & \epsilon_{l,k,n} = \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 + \sum_{(x,y) \neq (l,k)} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_y,y,n} \mvec{m}{x,y,n} \right |^2 \\ \nonumber
&\hspace{0.5cm}& \quad {} + N_0 \, \|\mvec{w}{l,k,n}\|^2 \IEEEyessubnumber \label{kkt-mse-1.2} \\
\sigma_{l,k,n} : & \hspace{0.5cm} & - \log(\tilde{\epsilon}_{l,k,n}) - \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\tilde{\epsilon}_{l,k,n}} = t_{l,k,n} \, \log(2) \IEEEyessubnumber \label{kkt-mse-1.3} \\
\delta_b : & \hspace{0.5cm} & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{kkt-mse-1.4}
\end{IEEEeqnarray}
where \me{\alpha_{l,k,n},\sigma_{l,k,n}} and \me{\delta_b} are the dual variables corresponding to the constraints defined in \eqref{kkt-mse-1.2}, \eqref{kkt-mse-1.3} and \eqref{kkt-mse-1.4}. The equality condition is used for \eqref{kkt-mse-1.2} and \eqref{kkt-mse-1.3}, since the variables \me{\epsilon_{l,k,n}} and \me{t_{l,k,n}} are used for the corresponding expressions on the other side.

Now the Lagrangian of the problem expressed in \eqref{kkt-mse-1}, with the corresponding dual variables, is used obtain the \ac{KKT} expressions by partially differentiating with respect to the variables \me{t_{l,k,n}}, \me{\epsilon_{l,k,n}}, \me{m_{l,k,n}}, \me{w_{l,k,n}} and the dual variables \me{\alpha_{l,k,n},\sigma_{l,k,n}} and \me{\delta_b}. Since the problem of precoder design using \ac{MSE} reformulation is solved by the iterative method, we use the iteration index \me{i} in the superscript to represent the variables at the respective iteration instants. The discussions on the \ac{KKT} expressions are provided in the  Appendix \ref{a-1}. Upon solving the \ac{KKT} expressions, we obtain the following iterative solution with the variables update as
\begin{IEEEeqnarray}{rCl} \label{kkt-mse-4}
\alpha^{(i)}_{l,k,n} &=& \dfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i-1)}}, \IEEEyessubnumber \label{kkt-mse-4.1} \\
\sigma_{l,k,n}^{(i)} &=& \dfrac{q \left [ a_k \, \left (Q_k - \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n}^{(i-1)} \right )^{(q-1)} \right ]}{\log(2)}, \IEEEyessubnumber \label{kkt-mse-4.2} \\
\mvec{m}{l,k,n}^{(i)} &=& \left ( \sum_{x \in \mc{U}} \sum_{y=1}^L \alpha_{y,x,n}^{(i)} \mvec{H}{b_k,x,n}^\herm \mvec{w}{y,x,n}^{(i-1)} \mvec{w}{y,x,n}^{\herm \, {(i-1)}} \mvec{H}{b_k,x,n} + \delta_b \mbf{I}_{N_T} \right )^{-1} \alpha^{(i)}_{l,k,n} \mvec{H}{b_k,k,n}^\herm \mvec{w}{l,k,n}^{(i-1)} \IEEEyessubnumber \label{kkt-mse-4.3} \\
\epsilon_{l,k,n}^{(i)} &=& \left | 1 - \mvec{w}{l,k,n}^{\herm \, (i-1)} \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}^{(i)} \right |^2 + \sum_{(x,y) \neq (l,k)} \left | \mvec{w}{l,k,n}^{\mathrm{H} \, (i-1)} \mvec{H}{b_y,y,n} \mvec{m}{x,y,n}^{(i)} \right |^2 \\ \nonumber
&+& \quad  N_0 \, \|\mvec{w}{l,k,n}^{(i-1)}\|^2,\quad \forall k \in \mc{U}, \forall n \in \mc{C} \IEEEyessubnumber \label{kkt-mse-4.4} \\
t_{l,k,n}^{(i)} &=&  -\log_2(\epsilon_{l,k,n}^{(i-1)}) - \frac{\left ( \epsilon_{l,k,n}^{(i)} - \epsilon_{l,k,n}^{(i-1)} \right ) }{\log(2) \, \epsilon_{l,k,n}^{(i-1)}}, \IEEEyessubnumber \label{kkt-mse-4.5} \\
\mvec{w}{l,k,n}^{(i)} &=& \left ( \sum_{(x,y) \neq (l,k)} \mvec{H}{b_y,k,n} \mvec{m}{x,y,n}^{(i-1)} \mvec{m}{x,y,n}^{\herm \, (i-1)} \mvec{H}{b_{y},k,n}^\herm + \mathbf{I}_{N_R} \right ) ^{-1} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n}^{(i-1)}, \IEEEyessubnumber \label{kkt-mse-4.6}
\end{IEEEeqnarray}
\me{\forall k \in \mc{U}}, \me{\forall n \in \set{1,\ldots,N}} and \me{\forall l \in \set{1,\ldots,L}}.

The \ac{KKT} solutions provided in \eqref{kkt-mse-4} are solved in an iterative manner by initializing the \ac{MSE} variables \me{\epsilon_{l,k,n}^{(0)}} and the throughput variables \me{t_{l,k,n}^{(0)}} randomly. It can be initiated by solving \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5} for a random transmit precoders \me{\mbf{M}_{k,n}} satisfying \eqref{kkt-mse-1.4}. It can be noted that the proposed iterative algorithm converges faster than the \acl{PD} and the \ac{ADMM} approach\footnote{from our simulations}, since the later schemes are based on the subgradient approach and the former one achieves the \ac{KKT} points at each iteration. In \eqref{kkt-mse-4}, all expressions provide the closed form solution for the respective variables except the transmit precoders in \eqref{kkt-mse-4.3}, which depends on the variable \me{\delta_b}. It can be seen that the variable \me{\delta_b} are associated only to the \ac{BS} \me{b}, which can be efficiently solved by the bisection method satisfying the power constraint \eqref{kkt-mse-1.4}. After each iteration instant, the transmit and the receive precoders are exchanged across the coordinating \acp{BS} in \me{\mc{B}} in order to reach the optimal point in the next iteration.

>>>>

By removing the absolute value operator from \eqref{eqn-mse-2}, the problem can be rewritten as
\begin{IEEEeqnarray}{rCl}\label{kkt-mse-1}
\underset{\substack{t_{l,k,n},\mvec{M}{k,n}, \alpha_{l,k,n}, \\ \delta_{b},\sigma_{l,k,n},\mvec{W}{k,n}}}{\text{minimize}} &\hspace{0.5cm}& \sum_{b \in \mc{B}} \sum_{k \in \mc{U}_b} \, a_k \, \left ( Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} t_{l,k,n} \right )^q \IEEEyessubnumber \label{kkt-mse-1.1} \\
\text{subject to} & \hspace{0.5cm} & \nonumber \\
\alpha_{l,k,n} : & \hspace{0.5cm} & \epsilon_{l,k,n} = \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 + \sum_{(x,y) \neq (l,k)} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_y,y,n} \mvec{m}{x,y,n} \right |^2 \\ \nonumber
&\hspace{0.5cm}& \quad {} + N_0 \, \|\mvec{w}{l,k,n}\|^2 \IEEEyessubnumber \label{kkt-mse-1.2} \\
\sigma_{l,k,n} : & \hspace{0.5cm} & - \log(\tilde{\epsilon}_{l,k,n}) - \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\tilde{\epsilon}_{l,k,n}} \geq t_{l,k,n} \, \log(2) \IEEEyessubnumber \label{kkt-mse-1.3} \\
\delta_b : & \hspace{0.5cm} & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{kkt-mse-1.4}
\end{IEEEeqnarray}
where \me{\alpha_{l,k,n},\sigma_{l,k,n}} and \me{\delta_b} are the dual variables corresponding to the constraints defined in \eqref{kkt-mse-1.2}, \eqref{kkt-mse-1.3} and \eqref{kkt-mse-1.4}. The equality condition is used for \eqref{kkt-mse-1.2} and \eqref{kkt-mse-1.3}, since the variables \me{\epsilon_{l,k,n}} and \me{t_{l,k,n}} are used for the corresponding expressions on the other side.

Now the Lagrangian of the problem expressed in \eqref{kkt-mse-1}, with the corresponding dual variables, is used obtain the \ac{KKT} expressions by partially differentiating with respect to the primal variables \me{t_{l,k,n}}, \me{\epsilon_{l,k,n}}, \me{m_{l,k,n}}, \me{w_{l,k,n}} and the dual variables \me{\alpha_{l,k,n},\sigma_{l,k,n}} and \me{\delta_b}. Since the precoder design is solved in an iterative fashion, the superscript \me{(i)} denotes the iteration instants. The detailed \ac{KKT} expressions are provided in the  Appendix \ref{a-1}. Upon solving the \ac{KKT} expressions, we obtain the following iterative solution as
\begin{IEEEeqnarray}{rCl} \label{kkt-mse-4}
\mvec{m}{l,k,n}^{(i)} &=& \left ( \sum_{x \in \mc{U}} \sum_{y=1}^L \alpha_{y,x,n}^{(i-1)} \mvec{H}{b_k,x,n}^\herm \mvec{w}{y,x,n}^{(i-1)} \mvec{w}{y,x,n}^{\herm \, {(i-1)}} \mvec{H}{b_k,x,n} + \delta_b \mbf{I}_{N_T} \right )^{-1} \alpha^{(i-1)}_{l,k,n} \mvec{H}{b_k,k,n}^\herm \mvec{w}{l,k,n}^{(i-1)} \IEEEyessubnumber \label{kkt-mse-4.3} \\
\mvec{w}{l,k,n}^{(i)} &=& \left ( \sum_{x\in\mc{U}}\sum_{y=1}^L \mvec{H}{b_x,k,n} \mvec{m}{y,x,n}^{(i)} \mvec{m}{y,x,n}^{\herm \, (i)} \mvec{H}{b_{x},k,n}^\herm + \mathbf{I}_{N_R} \right ) ^{-1} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n}^{(i)}, \IEEEyessubnumber \label{kkt-mse-4.6} \\
\epsilon_{l,k,n}^{(i)} &=& \left | 1 - \mvec{w}{l,k,n}^{\herm \, (i)} \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}^{(i)} \right |^2 + \sum_{(x,y) \neq (l,k)} \left | \mvec{w}{l,k,n}^{\mathrm{H} \, (i)} \mvec{H}{b_y,y,n} \mvec{m}{x,y,n}^{(i)} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}^{(i)}\|^2, \IEEEyessubnumber \label{kkt-mse-4.4} \\
t_{l,k,n}^{(i)} &=&  -\log_2(\epsilon_{l,k,n}^{(i-1)}) - \frac{\left ( \epsilon_{l,k,n}^{(i)} - \epsilon_{l,k,n}^{(i-1)} \right ) }{\log(2) \, \epsilon_{l,k,n}^{(i-1)}}, \IEEEyessubnumber \label{kkt-mse-4.5} \\
\sigma_{l,k,n}^{(i)} &=& q \,a_k \, \log_2(e)  \, \left (Q_k - \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n}^{(i)} \right )^{(q-1)}, \IEEEyessubnumber \label{kkt-mse-4.2} \\
\alpha^{(i)}_{l,k,n} &=& \alpha^{(i-1)}_{l,k,n} + \rho \left ( \frac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ), \IEEEyessubnumber \label{kkt-mse-4.1}
\end{IEEEeqnarray}
\me{\forall k \in \mc{U}}, \me{\forall n \in \set{1,\ldots,N}} and \me{\forall l \in \set{1,\ldots,L}}. The dual variable \me{\alpha^{(i)}} is updated with the memory to avoid any oscillations due to the over allocation of the resources. The step size \me{\rho} is used to reduce the effect on the precoder designs on the excess allocations.

The \ac{KKT} solutions provided in \eqref{kkt-mse-4} are solved in an iterative manner by initializing the transmit and the receive precoders \me{\mvec{M}{k,n},\mvec{W}{k,n}} with the single user beamforming and the \ac{MMSE} vectors. The dual variable \me{\alpha}'s corresponding to precoder weights are initialized with ones to provide equal priorities to all streams. With the initial values, the closed form solutions mentioned in \eqref{kkt-mse-4} are evaluated sequentially until convergence or to certain accuracy. In \eqref{kkt-mse-4}, all expressions are provided in a closed form except the transmit precoders \eqref{kkt-mse-4.3}, which depends on the \ac{BS} specific dual variable \me{\delta_b}. It can be solved efficiently by the bisection method satisfying the power constraint \eqref{kkt-mse-1.4}. After each iteration instant, the transmit and the receive precoders are exchanged across the coordinating \acp{BS} in \me{\mc{B}} to obtain the next operating point.

### 2014-07-10 Report/mse_formulation.tex
modified for KKT MMSE

In this section, we provide an alternative formulation for the \ac{QM} \ac{JSFRA} scheme using the \ac{MSE} equivalence with the capacity as in \cite{wmmse_shi,kaleva2012weighted} for the \ac{WSRM} problem. The proposed formulation uses the well known relation between the capacity and the \ac{MSE} given by \me{C(\gamma) = \log_2 \left ( {1}/{\epsilon} \right )}, where \me{\epsilon = {1} / {(1 + \gamma)}} is the \ac{MSE} expression for decoding the transmitted symbols using \ac{MMSE} receivers and \me{C} being the capacity of the system. Let \me{{v}^{\prime}_k = Q_k - \sum_{n=1}^N \sum_{l=1}^L t_{l,k,n}} denotes the queue deviation corresponding to the user \me{k} and \me{\tilde{{v}}^{\prime}_k \triangleq a_k^{1/{q}}v^{\prime}_k} represents the weighted equivalent. Now, using the \ac{MSE} relation, \eqref{eqn-6} is written as
\begin{IEEEeqnarray}{rCl}\label{eqn-mse-1}
\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n},\\\epsilon_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}^{\prime}  \|_q \IEEEyessubnumber \label{eqn-mse-1.1} \\
\text{subject to} & \quad & t_{l,k,n} \leq -\log_2(\epsilon_{l,k,n}), \IEEEyessubnumber \label{eqn-mse-1.2} \\
& \quad & \epsilon_{l,k,n} \geq  \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 + \sum_{(j,i) \neq (l,k)} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,i,n} \mvec{m}{j,i,n} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}\|^2 \IEEEyessubnumber \label{eqn-mse-1.3} \\
&\quad& \text{and} \; \eqref{eqn-4.3},  \IEEEyessubnumber \label{eqn-mse-1.4}
\end{IEEEeqnarray}
where \eqref{eqn-mse-1.3} bounds the \ac{MSE} by \me{\epsilon_{l,k,n}} and \eqref{eqn-mse-1.2} relaxes the transmitted rate \me{{t}_{l,k,n}} using the the \ac{MSE} relation.

>>>>

In this section, we use an alternative formulation for the \acl{QM} \ac{JSFRA} scheme using the \ac{MSE} equivalence with the capacity as in \cite{christensen2008weighted,wmmse_shi} for the \ac{WSRM} problem. The proposed formulation uses the well known relation between the capacity and the \ac{MSE} given by \me{C(\gamma) = \log_2 \left ( {1}/{\epsilon} \right )} \cite{viswanath1999optimal,mse_duality}, where \me{\epsilon = {1} / {(1 + \gamma)}} is the \ac{MSE} expression for decoding the transmitted symbols using \ac{MMSE} receivers and \me{C} being the capacity of the system. Let \me{{v}^{\prime}_k = Q_k - \sum_{n=1}^N \sum_{l=1}^L t_{l,k,n}} denotes the queue deviation corresponding to the user \me{k} and \me{\tilde{{v}}^{\prime}_k \triangleq a_k^{1/{q}}v^{\prime}_k} represents the weighted equivalent. Now, using the \ac{MSE} relation, \eqref{eqn-6} is written as
\begin{IEEEeqnarray}{rCl}\label{eqn-mse-1}
\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n},\\\epsilon_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}^{\prime}  \|_q \IEEEyessubnumber \label{eqn-mse-1.1} \\
\text{subject to} & \quad & t_{l,k,n} \leq -\log_2(\epsilon_{l,k,n}), \IEEEyessubnumber \label{eqn-mse-1.2} \\
& \quad & \epsilon_{l,k,n} \geq  \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 + \sum_{(j,i) \neq (l,k)} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,i,n} \mvec{m}{j,i,n} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}\|^2 \IEEEyessubnumber \label{eqn-mse-1.3} \\
&\quad& \text{and} \; \eqref{eqn-4.3},  \IEEEyessubnumber \label{eqn-mse-1.4}
\end{IEEEeqnarray}
where \eqref{eqn-mse-1.3} bounds the \ac{MSE} by \me{\epsilon_{l,k,n}} and \eqref{eqn-mse-1.2} relaxes the transmitted rate \me{{t}_{l,k,n}} using the the \ac{MSE} relation.

### 2014-07-10 Report/problem_formulation.tex
modified for KKT MMSE

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by the limiting it to a constant term. It can be seen that \eqref{eqn-4.2} provides similar expression when the exponent is set to be \me{q=2} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2.
\end{equation}
It can seen that \eqref{pf-1} is equivalent to the \eqref{backpreassure} algorithm when we ignore the second order terms from \eqref{pf-1}. It can be seen that by dropping the second order term from the \ac{Q-WSRM} formulation, we require an explicit rate constraint \eqref{eqn-3.1.4} to limit the allocations beyond the respective user requirements. In the current queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function itself. Since the objective is to minimize the number of queued packets, the sub-channels are coupled in this formulation as compared to the traditional \ac{WSRM} objective eventhough the interference from the adjacent sub-channels are zero (due to the \ac{OFDM} scheme). The queue deviation objective provides an alternate and more complete way to perform the resource allocation as compared to the \eqref{q_gen_sum} approach.

Before proceeding further, we note that the constraint in \eqref{eqn-4.2} is handled implicitly by the definition of the \me{q}-norm in the objective of \eqref{eqn-3}. As a proof, suppose that \me{t_k>Q_k} for a certain \me{k} at optimum, i.e., \me{-{v}_k=t_k-Q_k>0}. Then there exists \me{\delta_k>0} such that \me{-{v}^{\prime}_k=t^{\prime}_k-Q_k<-{v}_k} where \me{t^{\prime}_k=t_k-\delta_k}. Since \me{\|\tilde{\mbf{v}}\|_q=\| |\tilde{\mbf{v}}| \|_q=\||-\tilde{\mbf{v}}|\|_q}, this means that the newly created vector \me{\mbf{t}^{\prime}} achieves a smaller objective which contradicts with the fact that an optimal solution has been obtained. We comment on the choice of  the norm \me{q} on the objective as below \cite{berry2004cross,qps_cioffi}.
\begin{itemize}
\item With \me{q = 1}, the objective results in greedy allocation \textit{i.e}, emptying the queue of users with good channel condition before considering the users with worse channel conditions. As a special case, it is easy to see that \eqref{eqn-3} reduces to the \ac{WSRM} problem \eqref{gen_sum} when the queue size is large enough for all users.
\item With \me{q = 2}, the objective prioritizes users with higher number of queued packets before considering the users with a smaller number of backlogged packets. For example, it could be more ideal for the delay limited scenario when the packet arrival rates of the users are similar, since the backlogged packets is proportional to the delay in the transmission following the Little's law \cite{neely2010stochastic}.
\item With \(q = \infty \), the objective minimizes the maximum number of queued packets among users with the current transmission, thereby providing queue fairness by allocating the resources proportional to the number of backlogged packets.
\end{itemize}

>>>>

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by the limiting it to a constant value. Eq. \eqref{eqn-4.2} provides similar expression when the exponent is set to be \me{q=2} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2.
\end{equation}
It is evident that \eqref{pf-1} is equivalent to \eqref{backpreassure} when we ignore the second order terms from \eqref{pf-1}. Limiting \me{t_k^2} by a constant value, imposes the additional rate constraint \eqref{eqn-3.1.4} in the \ac{Q-WSRM} formulation to avoid the resource wastage in the form of over allocation. In the current queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function itself. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRM} problems include the sub-channels jointly to achive an efficient allocation by identifying the optimal space-frequency resource for each user in the system. The queue deviation objective provides an alternate and a more efficient way to perform the resource allocation as compared to the \eqref{q_gen_sum} approach.

Before proceeding further, we note that the constraint in \eqref{eqn-4.2} is handled implicitly by the definition of the \me{\ell_q} in the objective of \eqref{eqn-3}. As a proof, suppose that \me{t_k>Q_k} for a certain \me{k} at optimum, i.e., \me{-{v}_k=t_k-Q_k>0}. Then there exists \me{\delta_k>0} such that \me{-{v}^{\prime}_k=t^{\prime}_k-Q_k<-{v}_k} where \me{t^{\prime}_k=t_k-\delta_k}. Since \me{\|\tilde{\mbf{v}}\|_q=\| |\tilde{\mbf{v}}| \|_q=\||-\tilde{\mbf{v}}|\|_q}, this means that the newly created vector \me{\mbf{t}^{\prime}} achieves a smaller objective which contradicts with the fact that an optimal solution has been obtained. We comment on the choice of  the \me{\ell_q} on the objective as below \cite{berry2004cross,qps_cioffi}.
\begin{itemize}
\item With \me{q = 1}, the objective results in greedy allocation \textit{i.e}, emptying the queue of users with good channel condition before considering the users with worse channel conditions. As a special case, it is easy to see that \eqref{eqn-3} reduces to the \ac{WSRM} problem \eqref{gen_sum} when the queue size is large enough for all users.
\item With \me{q = 2}, the objective prioritizes users with higher number of queued packets before considering the users with a smaller number of backlogged packets. For example, it could be more ideal for the delay limited scenario when the packet arrival rates of the users are similar, since the backlogged packets is proportional to the delay in the transmission following the Little's law \cite{neely2010stochastic}.
\item With \(q = \infty \), the objective minimizes the maximum number of queued packets among users with the current transmission, thereby providing queue fairness by allocating the resources proportional to the number of backlogged packets.
\end{itemize}
